---
title: "Machine Learning Project"
output: html_document
---

# Summary

In this project we will take a data related to the quality of weight lifting performed by a group of people and predict the quality of weight lifting of another set of data.


# Data preparation

We first will load the data into a few tables then trim out the noise

```{r data_preparation}
require(data.table)
require(randomForest)
set.seed(1337)

# load libraries
library(caret)

# "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training_url <- "pml-training.csv"
# "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing_url <- "pml-testing.csv"
training <- read.csv(training_url, na.strings = c("NA", ""), strip.white=T)
testing <- read.csv(testing_url, na.strings = c("NA", ""), strip.white=T)
set_names <- names(training)
# summary(training)
# some worthless columns
worthless_column_indexes <- grep("X|user_name|.*timestamp.*", set_names)
training <- training[, -worthless_column_indexes]

# remove na item columns
na_items_index <- apply(training, 2, function(x) { sum(is.na(x)) })
training <- training[, which(na_items_index == 0)]

# remove columns that have low variability in order to train better
near_zero_variability_column_indexes <- nearZeroVar(training)
training <- training[, -near_zero_variability_column_indexes]

```

# Data partitioning

We split the data into two parts, a training and testing part.  For this part I did an 80/20 split between training and testing

```{r data_partitioning}

train_idx <- createDataPartition(training$classe, p=0.8, list=F)
t_training <- training[train_idx,]
t_testing <- training[-train_idx,]

```

# Training 

Now we'll create a random forrest training model for use on our partitioned test set.

```{r training }


# for performance, just run this the first time
if (!exists("train_model")){
  train_control <- trainControl(allowParallel=T, method="cv", number=4)
  train_model <- train(classe ~ ., data=t_training, model="rf", trControl=train_control)
}

train_model

train_model$finalModel

round(max(train_model$results$Accuracy), 4) * 100


```

Our Training model has an accuracy of 99.73%.

# Testing our model

Let's test our training model against our partitioned test set

```{r prediction_test}

# test out our training model on the testing portion
predict_model <- predict(train_model, newdata=t_testing)

# let's check out the confusion matrix
confusion_matrix <- confusionMatrix(t_testing$classe, predict_model)
confusion_matrix$table

# calculate our performance across the resample
postResample(predict_model, t_testing$classe)

```

In our testing model we got an accuracy of 99.8% which was extremely accurate

# Applying our model for the quiz

```{r quiz_apply}
predict_quiz <- predict(train_model, newdata=testing)

predict_quiz

```